##Loading and processing the data

```{r, echo=TRUE}
x <- read.csv("activity.csv")
colnames(x)
x$date <- as.Date(x$date)
```

##Mean total number of steps taken by day
##Total number of steps by day excluding missing values

```{r, echo=TRUE}
library(dplyr)
stepsbydate <- x  %>% 
  select(date, steps) %>%
  group_by(date) %>%
  summarise(stps= sum(steps)) %>%
  na.omit()
head(stepsbydate)
```
```{r}


##Histogram of total number of steps

```{r, echo=TRUE}
hist(stepsbydate$stps, xlab = "Total number of steps daily", main = "Histogram of total number of steps by day", breaks = 20)
```
##Mean and median number of steps taken by day
```{r, echo=TRUE}
mean(stepsbydate$stps)
median(stepsbydate$stps)
```

##Time series plot excluding NAs
```{r, echo=TRUE}
library(ggplot2)
xinterval <- x %>%
  select(interval, steps) %>%
  na.omit() %>%
  group_by(interval) %>%
  summarise(stps = mean(steps))
ggplot(xinterval, aes(x=interval, y=stps))+ geom_line() +
  xlab("5-minute interval")+
  ylab("average number of steps taken")
```

##The 5-min interval than on average contains the maximum number of steps

```{r, echo=TRUE}
xinterval[which(xinterval$stps==max(xinterval$stps)), ]
```

##Imputing missing values
##Calculate the total number of missing values
```{r, echo=TRUE}
NA_number <- length(which(is.na(x$steps)))
print(NA_number)
```
##Strategy and dataset 
```{r, echo=TRUE}
library(Hmisc)
ximp <- x
ximp$steps <- impute(x$steps, fun = mean)
```
##Histogram of total number of steps daily
```{r, echo=TRUE}
stepsbydateimp <- tapply(ximp$steps, ximp$date, sum)
qplot(stepsbydateimp, xlab = "Total steps by day imputed", ylab= "Frequency using binwith 500", binwidth=500)
```
##Calculate the mean and median
```{r, echo=TRUE}
mean(stepsbydateimp)
median(stepsbydateimp)
```
##Create a new factor variable in the dataset with two levels – “weekday” and “weekend”
```{r, echo =TRUE}
ximp$dateType <- ifelse(as.POSIXlt(ximp$date)$wday %in% c(0,6), "weekend", "weekday")
```
##Panel plot containing a time series plot
```{r, echo=TRUE}
avximp <- aggregate(steps ~ interval + dateType, data = ximp, mean)
ggplot(avximp, aes(interval, steps))+
  geom_line()+
  facet_grid(dateType ~ .) +
  xlab("5 minute interval") +
  ylab("average number of steps")
```


